---
title: "HW2 STA521 Fall18"
author: '[Your Name Here, netid and github username here]'
date: "Due September 19, 2017"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

## Backgound Reading

Readings: Chapters 3-4 in Weisberg Applied Linear Regression


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This exercise involves the UN data set from `alr3` package. Install `alr3` and the `car` packages and load the data to answer the following questions adding your code in the code chunks.  Please add appropriate code to the chunks to suppress messages and warnings as needed once you are sure the code is working properly and remove instructions if no longer needed. Figures should have informative captions. Please switch the output to pdf for your final version to upload to Sakai. **Remove these instructions for final submission**


## Exploratory Data Analysis

0.  Preliminary read in the data.  After testing, modify the code chunk so that output, messages and warnings are suppressed.  *Exclude text from final*

```{r data}
library(alr3)
data(UN3, package="alr3")
help(UN3) 
library(car)
```


1. Create a summary of the data.  How many variables have missing data?  Which are quantitative and which are qualtitative?

```{r}

```

2. What is the mean and standard deviation of each quantitative predictor?  Provide in a nicely formatted table.

```{r}

```


3. Investigate the predictors graphically, using scatterplots or other tools of your choice. Create some plots
highlighting the relationships among the predictors. Comment
on your findings regarding trying to predict `ModernC` from the other variables.  Are there potential outliers, nonlinear relationships or transformations that appear to be needed based on your graphical EDA?

```{r}

```

## Model Fitting

4.  Use the `lm()` function to perform a multiple linear regression with `ModernC` as the response and all other variables as the predictors, using the formula `ModernC ~ .`, where the `.` includes all remaining variables in the dataframe.  Create  diagnostic residual plot from the linear model object and comment on results regarding assumptions.  How many observations are used in your model fitting?

```{r}

```

5. Examine added variable plots `car::avPlot` or `car::avPlots`  for your model above. Are there any plots that suggest that transformations are needed for any of the terms in the model? Describe. Is it likely that any of the localities are influential for any of the terms?  Which localities?  Which terms?  

```{r}

```

6.  Using the Box-Tidwell  `car::boxTidwell` or graphical methods find appropriate transformations of the predictor variables to be used as predictors in the linear model.  If any predictors are negative, you may need to transform so that they are non-negative.  Describe your method and  the resulting transformations.


```{r}

```

7. Given the selected transformations of the predictors, select a transformation of the response using `MASS::boxcox` or `car::boxCox` and justify.


```{r}

```

8.  Fit the regression using the transformed variables.  Provide residual plots and added variables plots and comment.  If you feel that you need additional transformations, repeat any steps until you feel satisfied.


```{r}

```

9.  Are there any outliers or influential points in the data?  Explain.  If so, refit the model after removing any outliers and comment on residual plots.


```{r}

```

## Summary of Results

10. For your final model, provide summaries of coefficients with 95% confidence intervals in a nice table with interpretations of each coefficient.  These should be in terms of the original units! 


```{r}

```


11. Provide a paragraph summarizing your final model  and findings suitable for the US envoy to the UN after adjusting for outliers or influential points.   You should provide a justification for any case deletions in your final model


```{r}

```


## Methodology

12. We will use the United Nations data to verify properties of added-variable plots for the 'full' model 
$$E[\log(\text{Fertility}) \mid \log(\text{PPgdp}) = x_1, \text{Purban} = x_2) = \beta_0 + \beta_1 x_1 + \beta_2 x_2$$
There is nothing special about the two-predictor regression mean function but we are using it for simplicity. 

    a) Regress $\log(\text{Fertility})$ on $\text{Purban}$ and obtain the residuals; call them `e_Y` for example. 
```{r}

```

    b)  Regress $\log(\text{PPgdp})$ on $\text{Purban}$ and obtain the residuals; call them `e_X` for example. 
```{r}

```

    c) Create an added-variable scatter plot with `e_Y` on the y-axis and `e_X` on the x-axis.   Compare to the output of `avPlots` using the full model.  Are they the same?    

```{r}

```
    
    d) Prove that the intercept in the added variable scatter plot will always be zero.  _Hint:  use the fact that if $H$ is the project matrix which contains a column of ones, then $1_n^T (I - H) = 0$.  Use this to show that the sample mean of residuals will always be zero if there is an intercept._

    e) Fit a regression of `e_Y` on `e_X` and show that the estimated coefficient for $\log(\text{PPgdp})$ is the same as the estimated slope in the $\log(\text{PPgdp})$ in the full model with both predictors.This correctly suggests that all the estimates in a multiple linear regressionmodel are adjusted for all the other terms in the mean function.

```{r}

```

    f) show that the residuals in the added-variable plot are identical to the residuals from the mean function with both predictors.

```{r}

```

    g) Show that the t-test for the coefficient for $\log(\text{PPgdp})$ is not quite the same from the added-variable plot and from the regression with both terms, and explain why they are slightly different. 

```{r}

```

    h) Optional:  For multiple regression with more than 2 predictors, say a full model given by `Y ~ X1 + X2 + ... Xp`   we create the added variable plot for variable `j` by regressing `Y` on all of the `X`'s except `Xj` to form `e_Y` and then regressing `Xj` on all of the other X's to form `e_X`.  Confirm that the slope in the manually constructed added variable plot for one of the predictors  in Ex. 10 is the same as the estimate from your model. 
